---
title: "multi_sp_plots"
author: "Joe DeMarchi"
date: "2025-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Libraries
```{r}
library(lme4)
library(ggplot2)
library(ggpubr)
library(data.table)
library(rethinking)
library(dplyr)
library(cowplot)

setwd("C:/Users/demar/OneDrive/Documents/UTK/Research/Projects/multi_sp_competence")
```

### load Rds files
```{r}
growth_summary <- readRDS("sum_stats/growth_summary.rds")
growth_hpdi <- readRDS("sum_stats/growth_hpdi.rds")
growth_trajectory <- readRDS("sum_stats/growth_trajectory.rds")

transmission_summary <- readRDS("sum_stats/transmission_summary.rds")
transmission_hpdi  <- readRDS("sum_stats/transmission_hpdi.rds")
transmission_traj <- readRDS("sum_stats/transmission_traj.rds")

initial_infection_summary <- readRDS("sum_stats/initial_infection_summary.rds")
initial_infection_hpdi <- readRDS("sum_stats/initial_infection_hpdi.rds")

shedding_summary <- readRDS("sum_stats/shedding_summary.rds")
shedding_hpdi <- readRDS("sum_stats/shedding_hpdi.rds")

```

## load adult species data

```{r}
# all adult species
load("data/ipm_dat.RData")
# green frogs
load("data/green_ipm.Rdata")
# newts
load("data/newt_ipm.Rdata")
# bullfrog adults
load("data/bull_ipm.Rdata")

```



## Load tadpole species data
```{r}
btad <- fread("data/btad_master.csv")
```


### Setup tadpole data
```{r}
# setup data
# isolate mouth qty and shed

btad <- btad %>%
  mutate(
    dose_label = case_when(
      dose == 1e5 ~ "1x10^5",
      dose == 1e6 ~ "1x10^6",
      dose == 1e7 ~ "1x10^7",
      TRUE ~ NA_character_  # catch any unexpected values
    ),
    # numeric log10 dose level
    dose_level = log10(dose)
  )

# remove NA
btad <- na.omit(btad)
```

## Average infection load across species
```{r}
head(btad)
## plot trajectories
# 
# set up data including 0's
cols <- c("ID", "trt", "day", "swabZE", "species", "waterZE10")
dat <- rbind(
  green[, ..cols], 
  newt[, ..cols],
  bull[, ..cols]
) 

cols_tad <- c("id","dose", "mouth_qty", "shed_qty", "takedown_day")

dat <- dat %>%
  select(all_of(cols)) %>%
  filter(swabZE > 0) %>%
  mutate(
    logswab = log(swabZE)
  )

dat_tad <- btad %>%
  select(all_of(cols_tad)) %>%
  filter(mouth_qty > 0) %>% 
  mutate(
    species = "Tadpole",
    logswab = log(mouth_qty)
  )

mean_swab <- data.frame(
  species = c("R.catesbeianus", "R.clamitans", "N.viridescens", "Tadpole"),
  mean_logswab = c(
    mean(dat$logswab[dat$species == "R.catesbeianus"], na.rm = TRUE),
    mean(dat$logswab[dat$species == "R.clamitans"], na.rm = TRUE),
    mean(dat$logswab[dat$species == "N.viridescens"], na.rm = TRUE),
    mean(dat_tad$logswab, na.rm = TRUE)),
  logswab_se = c(
    sd(dat$logswab[dat$species == "R.catesbeianus"], na.rm = TRUE) / sqrt(sum(!is.na(dat$logswab[dat$species == "R.catesbeianus"]))),
    sd(dat$logswab[dat$species == "R.clamitans"], na.rm = TRUE) / sqrt(sum(!is.na(dat$logswab[dat$species == "R.clamitans"]))),
    sd(dat$logswab[dat$species == "N.viridescens"], na.rm = TRUE) / sqrt(sum(!is.na(dat$logswab[dat$species == "N.viridescens"]))),
    sd(dat_tad$logswab, na.rm = TRUE) / sqrt(nrow(dat_tad)
  ))
  )


mean_load_plot <- ggplot(mean_swab, aes(x = species, y = mean_logswab, fill = species)) +
  # Draw error bars first (behind)
  geom_errorbar(aes(ymin = mean_logswab - logswab_se, ymax = mean_logswab + logswab_se),
                width = 0.15, linewidth = 1, color = "gray23") +  # keep neutral outline
  # Draw points second (in front)
  geom_point(size = 5, shape = 21, stroke = 1.2, color = "gray23") +
  theme_classic() +
  scale_fill_manual(values = c(
    "R.catesbeianus" = "tan4", 
    "R.clamitans"   = "forestgreen", 
    "N.viridescens" = "darkorange3",
    "Tadpole"       = "aquamarine4"
  )) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = NULL, y = "log(load)")

avg_load_plt <- ggpar(mean_load_plot,
  ylab = "log(load)",
  font.y = c(20),
  font.tickslab = c(16),
  legend.position = "none"
)

ggsave("plots/avg_load_plot.png", avg_load_plt, width = 6, height = 7)

```


# Susceptibility/Transmission
```{r}

head(btad)
tdat <- btad
tdat$infec <- as.numeric(tdat$mouth_qty > 0)
names(tdat)
tdat$ln_dose <- log(tdat$dose)
tdat$dose10 <- log10(tdat$dose)


# Build model
tmod <- alist(
  infec ~ dbern(p),
  logit(p) <- b0  + b1*ln_dose,
  b0 ~ dnorm(0, 2),
  b1 ~ dnorm(0, 2)
)

## setup the rest of data
stan_data = tdat[,c("infec", "ln_dose")]

# Fit the model
tmod_fit = ulam(tmod, data = stan_data, iter = 5000, 
                warmup = 1000, chains = 4, cores = 4, log_lik = TRUE)

# Extract posterior samples

t_post <- extract.samples(tmod_fit)

```


## setup transmission data for plotting
```{r}
max_dose = log(1e7)
min_dose = log(1e4)

t_df = data.frame(species=rep(1, 100),
                  ln_dose=seq(min_dose, max_dose, length.out=100)) 

# calculate posteriors for each load
tres = link(tmod_fit, t_df)

#summarize
med_pred = apply(tres, 2, median)
lower = apply(tres, 2, quantile, 0.025)
upper = apply(tres, 2, quantile, 0.975)


#combine into l_df
t_df$pred = med_pred 
t_df$lower = lower 
t_df$upper = upper


#__________________________________________________________#

# Extract from new data posteriors to plot intercepts and slopes at mean dose

# mean dose is 5.5 so select two doses around mean
dose_1 = log(10^5.45)
dose_2 = log(10^5.55)
mean_dose = log(10^5.5)

#mean_dose_idx <- which.min(abs(as.numeric(t_df$ln_dose)))
# nearest values for targets
mean_dose_idx <- which.min(abs(t_df$ln_dose - mean_dose))
dose_1_idx    <- which.min(abs(t_df$ln_dose - dose_1))
dose_2_idx    <- which.min(abs(t_df$ln_dose - dose_2))

# extract posteriors
mean_res  <- tres[, mean_dose_idx, drop = FALSE] # p(dose = 5.5) posterior for each species
dose1_res <- tres[, dose_1_idx,    drop = FALSE] # p(dose = 5.45) posterior
dose2_res <- tres[, dose_2_idx,    drop = FALSE] # p(dose = 5.55) posterior for each species

# Find hpdi
sp_names <- c("Tadpole")


tint_hpdi <- HPDI(mean_res, prob = 0.95)
tint_df <- data.frame(
  species = sp_names,
  median = median(mean_res),
  #lower = unname(tint_hpdi[1]),
  #upper = unname(tint_hpdi[2]),
  lower = quantile(mean_res, 0.025),
  upper = quantile(mean_res, 0.975),
  row.names = NULL)

# long draws for violin
tint_plot <- data.frame(
  species = rep(sp_names, each = nrow(mean_res)),
  intercept = as.vector(mean_res)
)
#__________________________________________________________#
## Slope

# Use logit scale
logit_dose1_res <- log(dose1_res/ (1 - dose1_res)) # linearize on logit scale
logit_dose2_res <- log(dose2_res/ (1 - dose2_res)) #linearize on logit scale


#calculate slope for each posterior sample
slope_res <- (logit_dose2_res - logit_dose1_res) / (dose_2-dose_1)  # matrix with species as columns

#convert to data frame
tslope_res_df <- as.data.frame(slope_res)
colnames(tslope_res_df) <- c("Tadpole")

#summarize posterior slope distribution per species
t_slope <- data.frame(
  species = colnames(tslope_res_df),
  median = median(tslope_res_df$Tadpole),
  lower = apply(tslope_res_df, 2, quantile, 0.025),
  upper = apply(tslope_res_df, 2, quantile, 0.975)
)

## Calculate slope HPDI
tslope_hpdi <- apply(slope_res, 2, HPDI, prob = 0.95)
tslope_df <- data.frame(
  species = sp_names,
  median  = apply(slope_res, 2, median),
  lower   = tslope_hpdi[1,],
  upper   = tslope_hpdi[2,],
  row.names = NULL
)

tslope_draw <- data.frame(
  species = rep(sp_names, each = nrow(slope_res)),
  slope = as.vector(slope_res)
)


```

### Plot transmission results with all species

```{r}
# isolate intercepts
tran_sum <- transmission_summary[["intercept"]]
tall_int <- rbind(tran_sum, tint_df)

# load violins
  
## Plot by mean dose
t_int_plot <- ggplot(tall_int, aes(x = species, y = median, color = species)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") +  # Add error bars
  geom_point(size = 5,
             shape = 21,
             stroke = 0.5,
             aes(fill = species), color = "black") +  # Add points
  geom_violin(data = tint_plot, 
              aes(x = species, y = intercept, fill = species),
              alpha = 0.3,
              trim = FALSE,
              width = 0.8,
              inherit.aes = FALSE) +  # Add violin for tadpole
  geom_violin(data = transmission_hpdi[["intercept"]], 
              aes(x = species, y = intercept, fill = species),
              alpha = 0.3,
              trim = FALSE,
              width = 0.8,
              inherit.aes = FALSE) +  # Add violin for other species
  theme_classic() + 
  labs(
    x = NULL,
    y = "Infection probability"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "none"  # Remove legend since species are on x-axis
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) +  # Fill for points
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4"))   # Color for legend

### ________________________________________________________________________##
# Slopes

tran_slope <- transmission_summary[["slope"]]
tall_slope <- rbind(tran_slope, t_slope)

# plot slopes

t_slope_plot <- ggplot(tall_slope, aes(x = species, y = median, color = species)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") +  # Add error bars
  geom_point(size = 5,
             shape = 21,
             stroke = 0.5,
             aes(fill = species), color = "black") +  # Add points
  geom_violin(data = tslope_draw, 
              aes(x = species, y = slope, fill = species),
              alpha = 0.3,
              trim = FALSE,
              width = 0.8,
              inherit.aes = FALSE) +  # Add violin for tadpole
  geom_violin(data = transmission_hpdi[["slope"]], 
              aes(x = species, y = slope, fill = species),
              alpha = 0.3,
              trim = FALSE,
              width = 0.8,
              inherit.aes = FALSE) +  # Add violin for other species
  theme_classic() + 
  labs(
    x = NULL,
    y ="Effect of dose on \n\ infection probability \n\ (logit scale)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "none"  # Remove legend since species are on x-axis
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) +  # Fill for points
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4"))   # Color for legend

###________________________________________________________________________###
# Fix legends

#make the axis look good
t_int_plt <- ggpar(t_int_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)

#Fix slope axis
t_slope_plt <- ggpar(t_slope_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)
######______________________________________________________________________####
### Full trajectory plot
transmission_traj
t_df$species <- "Tadpole"
t_df$dose <- t_df$ln_dose/log(10)
transmission_traj$dose <- transmission_traj$dose/log(10)

tplot <- ggplot() +
  geom_line(data = t_df, aes(x = dose, y = pred, color = species), linewidth = 1) +
  geom_ribbon(data = t_df, aes(x = dose, ymin = lower, ymax = upper, fill = species), alpha = 0.1) +
  geom_line(data = transmission_traj, aes(x = dose, y = pred, color = species), linewidth = 1) +
  geom_ribbon(data = transmission_traj, aes(x = dose, ymin = lower, ymax = upper, fill = species), alpha = 0.1) +
  #geom_jitter(aes(x = dose, y = infec, color = species1), alpha = 0.2, height = 0, width = 0.1) +
  theme_classic() + 
  xlab("Exposure dose 1x10^ \n\ zoospores/mL") + ylab("Infection probability") + scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4")) +
    scale_fill_manual(values = c(
    "R.catesbeianus" = "tan4", 
    "R.clamitans" = "forestgreen", 
    "N.viridescens" = "darkorange3",
    "Tadpole" = "aquamarine4")
  ) +
      guides(color = guide_legend(title = "Species"),
         fill = guide_legend(title = "Species")) +
  theme(axis.ticks.length = unit(-0.15, "cm"))
 


tplt <- ggpar(tplot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14),
  legend.title = "Species",
  font.legend = c(14),
  
)

ggsave("plots/transmission_dose_response.png", tplt, width = 6, height = 5)

```

### Combine plots
```{r}
#combine plots
transmission_plt <- (t_int_plt | t_slope_plt) / tplt +
  plot_layout(
    widths = c(1, 1),   
    heights = c(1, 1)
  ) + 
    plot_annotation(tag_levels = "A"
  )
ggsave("plots/trans_tad_plots.png", transmission_plt, width = 8, height = 8)
```



### Growth
```{r}

# filter data and log load
gtad <- btad %>%
  filter(mouth_qty > 0) %>%
  mutate(
    load = log(mouth_qty),
    day  = takedown_day
  )

# create time level matrix
time_levels <- sort(unique(gtad$day))
T <- length(time_levels)

# combine matrix to take down day
gtad <- gtad %>%
  mutate(
    t_index = match(day, time_levels)  
  )


# set up for stan
stan_dat <- list(
  N = nrow(gtad),
  T = T,
  t_index = gtad$t_index,
  y = gtad$load
)


grow_mod <- cmdstan_model("growth_tadpole.stan")

fit <- grow_mod$sample(
  data = stan_dat,
  chains = 4, iter_warmup = 1000, iter_sampling = 5000, parallel_chains = 4
)

posterior <- fit$draws(c("a", "b", "sigma", "m", "v"))

fit$summary(c("a","b","sigma","m","v"))

growth_post <- as_draws_df(fit$draws())
```

### Extract growth posteriors

```{r}
growth_post <- as_draws_df(fit$draws())

# posterior draws (ndraws-length vectors)
a_draw <- growth_post$a
b_draw <- growth_post$b
#sig_draw <- growth_post$sigma  # only needed if you want yrep

# grid of loads
load_grid <- seq(min(gtad$load), max(gtad$load), length.out = 300)

# mu[draw, x]  (ndraws x 300)
mu <- outer(a_draw, rep(1, length(load_grid))) +
      outer(b_draw, load_grid)

# summarize per x (columns)
pred_df <- tibble(
  load  = load_grid,
  pred  = apply(mu, 2, median),
  lower = apply(mu, 2, quantile, 0.025),
  upper = apply(mu, 2, quantile, 0.975)
)

mean_load <- 3.70

# Posterior estimates plots
species <- c("Tadpole")

## slope

g_tad_slope2 <- data.table(
  species = species,
  slope = median(b_draw),
  slope_lower = quantile(b_draw, 0.025),
  slope_upper = quantile(b_draw, 0.975)
)

sp_names = c("Tadpole")


gtint_hpdi2 <- HPDI(a_draw, prob = 0.95)
gint_plot2 <- data.frame(
  species = rep(sp_names, length(a_draw)),
  intercept = as.vector(a_draw)
)

# update gtad_dist:

gtad_dist2 <- data.table(
  species = species,
  intercept = median(a_draw),
  intercept_lower = unname(gtint_hpdi2[1]),
  intercept_upper = unname(gtint_hpdi2[2])
)


gslope_plot2 <- data.frame(
  species = rep(sp_names, length(b_draw)),
  slope = as.vector(b_draw)
)


gtad_sum <- cbind(gtad_dist2, g_tad_slope2[, -1, with = FALSE])



```

### Growth trajectory plot

```{r}
gdat <- growth_trajectory
pred_df$species1 <- "Tadpole"
pred_df$species <- 4
pred_df$loadNext <- NA  # placeholder for loadNext

# find shared column names
common_cols <- intersect(names(gdat), names(pred_df))

# restrict both data frames
g_traj <- gdat %>% dplyr::select(all_of(common_cols))
tad_traj <- pred_df %>% dplyr::select(all_of(common_cols))

# combine
g_df <- bind_rows(g_traj, tad_traj)

gplot <- ggplot(g_df) +
  geom_point(aes(x = load, y = loadNext, color = species1), alpha = 0.2) +
 # geom_line(aes(x = lnload, y = lnwaterload, color = species1), alpha = 0.4) +
  geom_line(aes(x = load, y = pred, color = species1), linewidth = 1.05) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed") +
  geom_ribbon(aes(x = load, ymin = lower, ymax = upper, fill = species1), alpha = 0.1) +
  theme_classic() + xlab("log(load)") + ylab("log(loadNext)") + scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4")) +
    scale_fill_manual(values = c(
    "R.catesbeianus" = "tan4", 
    "R.clamitans" = "forestgreen", 
    "N.viridescens" = "darkorange3",
     "Tadpole" = "aquamarine4"
  )) +
  theme(axis.ticks.length = unit(-0.15, "cm"))

#make the plot look better
gplt <- ggpar(gplot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14),
  legend.title = "Species",
  font.legend = c(14),
  
)



```


### Growth plot

```{r}
## Intercept
gall_int <- growth_summary
tgall_int <- rbind(gall_int, gtad_sum)


g_int_plot <- ggplot(tgall_int, aes(x = species, y = intercept, color = species)) +
  geom_errorbar(aes(ymin = intercept_lower, ymax = intercept_upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") + 
  geom_point(aes(fill = species), 
             color = "black",
             size = 5,
             shape = 21,
             stroke = 0.5) +  
  geom_violin(data = gint_plot2,
              aes(x = species, y = intercept, fill = species),
              alpha = 0.3,
              trim = FALSE,
              inherit.aes = FALSE) +
  geom_violin(data = growth_hpdi[["intercept"]],
              aes(x = species, y = int_violin, fill = species),
              alpha = 0.3,
              trim = FALSE,
              inherit.aes = FALSE) +
  theme_classic() + 
  labs(
    x = NULL,
    y = "log(Bd growth rate)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) +  
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4"))   

##____________________________________________________________________________##

### slope


g_slope_plot <- ggplot(tgall_int, aes(x = species, y = slope, color = species)) +
  geom_errorbar(aes(ymin = slope_lower, ymax = slope_upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") +
  geom_point(size = 5,
             shape = 21,
             stroke = 0.5,
             aes(fill = species), color = "black") + 
  geom_violin(data = gslope_plot2,
              aes(x = species, y = slope, fill = species),
              alpha = 0.3,
              trim = FALSE,
              inherit.aes = FALSE) +
  geom_violin(data = growth_hpdi[["slope"]],
              aes(x = species, y = slope, fill = species),
              alpha = 0.3,
              trim = FALSE,
              inherit.aes = FALSE) +  # add violin of of other species
  theme_classic() + 
  labs(
    x = NULL,
    y ="Effect of load on \n\ growth \n\ (log zoospores)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # rotate x-axis labels
    legend.position = "none"  # remove legend
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) + 
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4")) 

###________________________________________________________________________##
# Fix plots

 # fix up intercept plot
g_int_plt <- ggpar(g_int_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)
#fix up slope plot
g_slope_plt <- ggpar(g_slope_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)


```
### Combine growth plots

```{r}
library(patchwork)

growth_plt <- (g_int_plt | g_slope_plt) / gplt +
  plot_layout(
    widths = c(1, 1),   
    heights = c(1, 1)
  ) + 
    plot_annotation(tag_levels = "A"
  )


# save growth plot
ggsave("plots/growth_tad_plots_multi.png", growth_plt, width = 8, height = 8)

```



# Shedding

```{r}
tad = fread("data/btad_master.csv")
## needed columns
cols <- c("id", "dose", "mouth_qty", "shed_qty")
# Drop the instances where
# Newts
trun_tad_dat <- tad[!(shed_qty == 0 & mouth_qty == 0) & !is.na(shed_qty) & !is.na(mouth_qty), ..cols]

shedding_mod = cmdstan_model("shedding_model1.stan")
# create species index
trun_tad_dat$species <- coerce_index(as.factor(trun_tad_dat$species))

# prepare data for inclusion in model
N = nrow(trun_tad_dat)
M = sum(trun_tad_dat$mouth_qty == 0) # Missing swabs
P = sum(trun_tad_dat$shed_qty == 0)

# Missing load
missing_ind = rep(0, N)
missing_ind[which(trun_tad_dat$mouth_qty == 0)] = 1:M

#Missing shed
missing_shed_ind = rep(0, N)
missing_shed_ind[which(trun_tad_dat$shed_qty == 0)] = 1:P

# rename and transform columns to fit into stan model
stan_data = list(N=N,
                 M=M,
                 P=P,
                 load=trun_tad_dat$mouth_qty,
                 shedding=trun_tad_dat$shed_qty,
                 missing_ind=missing_ind,
                 missing_shed_ind=missing_shed_ind)

#fit the model
tad_fit = shedding_mod$sample(data = stan_data,
                                    chains = 4,
                                    iter_warmup = 1000,
                                    iter_sampling = 5000,
                                    parallel_chains = 4)


tad_post <- as_draws_df(tad_fit$draws())

# Extract posterior samples for species-specific parameters
beta0 <- as.matrix(tad_post[, grep("beta0", colnames(tad_post))])
beta1 <- as.matrix(tad_post[, grep("beta1", colnames(tad_post))])

## make a function to pull out mean and confidence intervals

post_stats <- function(post_matrix) {
  apply(post_matrix, 2, function(col){
    c(median = mean(col),
      lower = quantile(col, 0.025),
      upper = quantile (col, 0.975))
  })
}

#calculate the summary statistics
beta0_tad_sum <- post_stats(beta0)
beta1_tad_sum <- post_stats(beta1)


# Generate posterior predictions for each species
new_vals <- seq(-3, 12, len = 100)          # Range for predictions
# Posterior predictions: matrix of [iterations x new_vals]
post_pred <- sapply(1:length(beta0), function(i) {
  beta0[i] + beta1[i] * new_vals
})

# Each column is one value of new_vals — take medians by row
med <- apply(post_pred, 1, median)
lower <- apply(post_pred, 1, quantile, probs = 0.025)
upper <- apply(post_pred, 1, quantile, probs = 0.975)

# create data frame
pred_tad_stats <- data.frame(
  new_vals = new_vals,
  median = med,
  lower = lower,
  upper = upper,
  species1 = rep("Tadpole", each = length(new_vals)))


# make intercept and slope data.frame

#make into a data frame
post_tad_res <- data.frame(
  species = c("Tadpole"),
  beta0_med = beta0_tad_sum["median", ],
  beta0_lower = beta0_tad_sum["lower.2.5%", ],
  beta0_upper = beta0_tad_sum["upper.97.5%", ],
  beta1_med = beta1_tad_sum["median", ],
  beta1_lower = beta1_tad_sum["lower.2.5%", ],
  beta1_upper = beta1_tad_sum["upper.97.5%", ]
)

species = c("Tadpole")

## Calculate long hpdi posterior distribution for plotting

b0_plot <- data.frame(
  species = rep(species, each = nrow(beta0)),
  intercept = as.vector(beta0)
)

b1_plot <- data.frame(
  species = rep(species, each = nrow(beta1)),
  slope = as.vector(beta1)
)


```


### Shedding plots

```{r}
## load multi_sp data
sp_post_res <- shedding_summary

post_res_plot <- rbind(sp_post_res, post_tad_res)

# intercept plot
shed_int_plot <- ggplot(post_res_plot, aes(x = species, y = beta0_med, color = species)) +
  geom_errorbar(aes(ymin = beta0_lower, ymax = beta0_upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") +  # Add error bars
  geom_point(size = 5,
             shape = 21,
             stroke = 0.5,
             aes(fill = species), color = "black") +  # Add points
  geom_violin(data = b0_plot, aes(x = species, y = intercept, fill = species),
              trim = FALSE,
              alpha = 0.3,
              inherit.aes = FALSE) +
  geom_violin(data = shedding_hpdi[["intercept"]], aes(x = species, y = intercept, fill = species),
              trim = FALSE,
              alpha = 0.3,
              inherit.aes = FALSE) +
  theme_classic() + 
  labs(
    x = NULL,
    y = "log(Bd shedding rate)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "none",  # Remove legend since species are on x-axis
    axis.ticks.length = unit(-0.15, "cm")
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) +  # Fill for points
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4"))  

#_______________________________________________________________________#

# fix intercept plot
shed_int_plt<- ggpar(shed_int_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)

#____________________________________________________________________________#
## setup data

### Slopes
shed_slope_plot <- ggplot(post_res_plot, aes(x = species, y = beta1_med, color = species)) +
  geom_errorbar(aes(ymin = beta1_lower, ymax = beta1_upper), 
                width = 0.2,
                linewidth = 1,
                color = "black") +  # Add error bars
  geom_point(size = 5,
             shape = 21,
             stroke = 0.5,
             aes(fill = species), color = "black") +  # Add points
  geom_violin(data = b1_plot, aes(x = species, y = slope, fill = species),
              trim = FALSE,
              alpha = 0.3,
              inherit.aes = FALSE) +
  geom_violin(data = shedding_hpdi[["slope"]], aes(x = species, y = slope, fill = species),
              trim = FALSE,
              alpha = 0.3,
              inherit.aes = FALSE) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "tomato3") +  # Add line at 0)
  theme_classic() + 
  labs(
    x = NULL,
    y = "Effect of load on \n\ Bd shedding rate"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.position = "none",  # Remove legend since species are on x-axis
    axis.ticks.length = unit(-0.15, "cm")
  ) +
  scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                               "R.clamitans" = "forestgreen", 
                               "N.viridescens" = "darkorange3",
                               "Tadpole" = "aquamarine4")) +  # Fill for points
  scale_color_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "N.viridescens" = "darkorange3",
                                "Tadpole" = "aquamarine4"))  



#____________________________________________________________________________#

# fix slope plot
shed_slope_plt <- ggpar(shed_slope_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
  
)


# Plot
# set up abline from the mean
ab <- 4.044998 #mean from multi_species work

# setup data

shed_dat <- readRDS("sum_stats/shedding_multi_sp.rds")
shed_pred <- readRDS("sum_stats/shedding_post_pred_stats.rds")

trun_tad_dat[, species := as.character(species)]
trun_tad_dat[, species := 4]

setnames(trun_tad_dat, old = c("mouth_qty", "shed_qty"),
             new = c("swabZE",   "waterZE10"))

# keep only these columns
tad_dat <- trun_tad_dat[, .(swabZE, waterZE10, species)]

sp_shed_dat <- shed_dat[, .(swabZE, waterZE10, species)]

shed_all <- rbind(sp_shed_dat, tad_dat)

pred_tad_stats$species <- "4"

pred_all <- rbind(shed_pred, pred_tad_stats)


shed_plot <- ggplot() +
  geom_point(data = shed_all[shed_all$swabZE != 0 & shed_all$waterZE10 != 0, ],
             aes(x = log(swabZE), y = log(waterZE10), color = as.factor(species)), alpha = 0.3) +
  # Species-specific lines
  geom_line(data = pred_all, aes(x = new_vals, y = median, color = species), linewidth = 1.15) +
  # Ribbons for confidence intervals
  geom_ribbon(data = pred_all, aes(x = new_vals, ymin = lower, ymax = upper, fill = species), alpha = 0.2) +
  # Add abline for each species plot
  geom_abline(slope = 1, intercept = ab, color = "tomato3", linetype = "dashed", linewidth = 0.8) +
  # Axis labels
  labs(x = "log(load)", 
       y = "log(shed)", 
       color = "Species", 
       fill = "Species", 
       group = "Point Type") +
  # Scales
  scale_color_manual(values = c("1" = "darkorange3", "2" = "tan4", "3" = "forestgreen", "4" = "aquamarine4"),
                     labels = c("1" = "N.viridescens", "2" = "R.catesbeianus", "3" = "R.clamitans",
                                "4" = "Tadpole")) +
  scale_fill_manual(values = c("1" = "darkorange3", "2" = "tan4", "3" = "forestgreen", "4" = "aquamarine4"),
                    labels = c("1" = "N.viridescens", "2" = "R.catesbeianus", "3" = "R.clamitans",
                               "4" = "Tadpole")) +
  facet_wrap(~species, labeller = labeller(species = c(
    "1" = "N.viridescens",
    "2" = "R.catesbeianus",
    "3" = "R.clamitans",
    "4" = "Tadpole"
  ))) +
  theme_classic() +
   theme(
    strip.text = element_text(size = 14, face = "italic"),
                              legend.position = "none",
    axis.ticks.length = unit(-0.15, "cm")
  )


#____________________________________________________________________________#

# fix plot axis
shed_plt <- ggpar(shed_plot,
  font.x = c(16),
  font.y = c(16),
  font.tickslab = c(14)
)

##_________________________________________________________________________##
#Full trajectory plot




```

## combine all the plots together

```{r}
shed_all <- (shed_int_plt | shed_slope_plt) / shed_plt +
                plot_layout(
    widths = c(1, 1),   
    heights = c(1, 1)
  ) + 
    plot_annotation(tag_levels = "A"
  )

ggsave("plots/shed_tad_plots.png", shed_all, width = 8, height = 10)
```

