---
title: "Q1_function"
author: "Joe DeMarchi"
date: "2025-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# set working directory
setwd("C:/Users/demar/OneDrive/Documents/UTK/Research/Projects/multi_sp_competence")


# Load required packages
library(dplyr)
library(ggplot2)
library(scales)
library(readr)
library(data.table)
```

Load Data
```{r}
source("multi_sp_fxns.R")

# load rds files
# load sp data
bull <- readRDS("data/bull_dist.rds")
green <- readRDS("data/green_dist.rds")
tad <- readRDS("data/btad_dist.rds")
newt <- readRDS("data/newt_dist.rds")


abundance = read.csv("amphibianTN_field/abundance.csv") #Site and species-level abundance
#abundance$species = factor(abundance$species, levels = species_levels)

abundance_wh <- subset(abundance, site == 15) # Subset for WH site)
abundance_wh <- subset(abundance_wh, species %in% c("novi", "racl", "raca"))

mean_abundance <- abundance_wh %>%
  group_by(species) %>%
  summarize(mean_abundance = mean(abund, na.rm = TRUE),
            sd_abundance   = sd(abund, na.rm = TRUE),
            n              = sum(!is.na(abund)),
            se_abundance   = sd_abundance / sqrt(n),
            min_abundance  = min(abund, na.rm = TRUE),
            max_abundance  = max(abund, na.rm = TRUE),
            .groups = "drop"
  )

mean_abundance


movement_distance <- data.frame(
  species = c("N.viridescens", "R.clamitans", "R.catesbeianus"),
  movement_distance = c(10, 40, 30)*0.5 # movement distances in m^3 per 6 days in a pond 0.5 m deep
)
```

## Q1. Under what ecological conditions is indirect transmission via a shared environmental zoospore pool sufficient to support Bd persistence?

### Set parameters
```{r}
source("multi_sp_fxns_q1.R")

# add in extra parameters needed
# Load rds files for other parameters
bull_par <- readRDS("params/bull_demo_pars.rds")
green_par <- readRDS("params/green_demo_pars.rds")
newt_par <- readRDS("params/newt_demo_pars.rds")

## 
# Green ADULTS ____________________________________________________________
# survival set to 1
steps_per_year <- round(365/6) # survives about 5 years
s_year <- 0.60
green$s0 <- (s_year)^(1/steps_per_year)
# Movment distance by 6 day time step (m2)
# converted to movement (m3) in build_transition_matrix function
green$movement_distance <- 40
## lower movement bound 20-30
## upper movement bound 60-80
# Equilibrium density
green$df_density <- 238
### Metamorphosis success
meta_years_mean <- 1.5
green$meta_successG <-  1 - exp(-(1/meta_years_mean)/steps_per_year)

##
# Bullfrog ADULTS ____________________________________________________________
s_year <- 0.70
steps_per_year <- round(365/6)
bull$s0 <- (s_year)^(1/steps_per_year)
# Movment distance by 6 day time step (m2)
# converted to movement (m3) in build_transition_matrix function
bull$movement_distance <- 30
## lower movement bound: 20-30
## upper movement bound 60-80
# bullfrog equilibrium density
bull$df_density <- 23 # 20

## Metamorphosis Success
meta_years_mean = 2.5
bull$meta_successB <- 1 - exp(-(1/meta_years_mean)/steps_per_year)

## Newt ADULTS ____________________________________________________________
s_year <- 0.50 # adult newts can live up to 13 years but average around 3 so similar to green frogs lets assume 7-9 and
steps_per_year <- round(365/6)
newt$s0 <- (s_year)^(1/steps_per_year)
newt$movement_distance <- 10 # avg 6 day movement
# newt equilibrium density
newt$df_density <- 872

#___________________________________________________________________________

# take mean of all parameters
newt_dist <- lapply(newt, function(x) median(x))
green_dist <- lapply(green, function(x) median(x))
bull_dist <- lapply(bull, function(x) median(x))

## Bullfrog tadpoles ___________________________________________________________
btad <- tad
# survival set to 1
s_year <- 0.15
steps_per_year <- round(365/6)
sb <- s_year^(1/steps_per_year)
btad$s0 <- sb
btad$movement_distance <- 6
# add loss parameters to tadpole params
btad$loss_int = bull_dist$loss_int
btad$loss_slope = bull_dist$loss_slope
btad$df_density <- 450
btad$meta_successB <- bull$meta_successB
btad$K_b <- bull_par$K_b
btad$emig_meta <- bull_par$emig_meta


## Green frog tadpoles ___________________________________________________________
gtad <- tad
# survival set to 1
s_year <- 0.10
steps_per_year <- round(365/6)
sg <- s_year^(1/steps_per_year)
gtad$s0 <- sg
gtad$movement_distance <- 3
# add loss parameters to tadpole params
gtad$loss_int = green_dist$loss_int
gtad$loss_slope = green_dist$loss_slope
gtad$df_density <- 1560
gtad$meta_successG <- green$meta_successG
gtad$K_g <- green_par$K_g
gtad$emig_meta <- green_par$emig_meta

# take mean of all parameters
btad_dist <- lapply(btad, function(x) median(x))
gtad_dist <- lapply(gtad, function(x) median(x))

## combine into one list
ipm_dists <- list(
  R.catesbeianus = modifyList(bull_dist,  bull_par),
  R.clamitans   = modifyList(green_dist, green_par),
  N.viridescens = modifyList(newt_dist,  newt_par),
  Raca.tadpole = btad_dist,
  Racl.tadpole = gtad_dist
)

# Save ipm dists
save(ipm_dists, file = "ipm_dists.Rdata")

```

### ADD in field data
```{r}
# Species and site labels ----

labels = c("hych"="H. chysoscelis","novi"="N. viridescens","pscr"="P. crucifer","psfe"="P. feriarum","raca"="R. catesbeianus","racl"="R. clamitans")
species_levels = names(labels)

# Set working directory ----
setwd("C:/Users/demar/OneDrive/Documents/UTK/Research/Projects/multi_sp_competence/amphibianTN_field")
# Load datasets ----


### Site & Survey level environmental data ----
# Read in dataset
pond_data = read.csv("pond_data.csv", stringsAsFactors = FALSE) #Pond sampling level data (may have multiple rows for same survey when multiple water quality measurements (ysi) were taken for a single site)

# Read in field descriptions
comments_df <- read.table("pond_data_comments.txt", header = TRUE, sep = "\t", stringsAsFactors = FALSE)

# Reattach comments
for (i in seq_len(nrow(comments_df))) {
  col_name <- comments_df$column[i]
  col_comment <- comments_df$comment[i]
  if (col_name %in% names(pond_data)) {
    comment(pond_data[[col_name]]) <- col_comment
  }
}

str(pond_data)

pond_data_lim = pond_data[pond_data$site_id==2 | pond_data$site_id==7 | pond_data$site_id==10 | pond_data$site_id==15,]


### Area ----
pond_area = read.csv("pond_area_df.csv") #Sampling-level data of pond area for Lotus and William Hastie
site_area = read.csv("site_area_df.csv") #Area of full site sampled (land and water)

### Water temperature ----
temperature = read.csv("temperature_dat.csv") #Site-level water temperature 

### Abundance ----
abundance = read.csv("abundance.csv") #Site and species-level abundance
abundance$species = factor(abundance$species, levels = species_levels)


## Interpolated data ----
### Water temperature ----
temperature_gam = read.csv("temp_gam_df.csv") #Site-level water temperature interpolated for the full year

### Abundance ----
abundance_gam = read.csv("abund_gam_df.csv") #Site and species-level abundance interpolated for the full year
abundance_gam$species = factor(abundance_gam$species, levels = species_levels)


abundance$species = factor(abundance$species, levels = species_levels)

### Calculate tadpole
#load data
tadpole_abundance<-read.csv("amphibianTN_field/pipe_sampling.csv")

# subset jsut for william hastie
tad_abund <- subset(tadpole_abundance, subsite_id %like% "WH")
head(tad_abund)
#find mean abundance for each species
# Individual species abundance

# mean tadpoles per pipe
mean_count <- mean(tad_abund$tadpole_count, na.rm = TRUE)

# pipe area and pond area
pipe_area <- 0.8  # m^2
pond_radius <- 15 # m
pond_area <- pi * pond_radius^2

# by night
night_density <- tadpole_abundance %>%
  group_by(survey_id) %>% 
  summarize(
    mean_count = mean(tadpole_count),
    density = mean_count / 0.8
  )

# Night abundance estimate
night_density$abund_est <- night_density$density*pond_area

mean(night_density$abund_est) # approx 1000 tadpoles


### boot strap it to check
boot_abund <- replicate(5000, {
  m <- mean(sample(night_density$density, replace = TRUE))
  m * pond_area
})

quantile(boot_abund, c(0.025, 0.5, 0.975)) # ~2000

## Figure out what proportion of total tadpoles are bullfrogs vs green frogs
# adult density
N_g <- 238
N_b <- 23

#clutch size
c_g <- 5000
c_b <- 15000
# surval of egg to tadpole
s = 0.10 # survival rate or eggs

#total average tadpoles in pond

total_tads <- 2000

## find out proportion of tadpoles

## Relative contribution (egg × clutch × survival)
T_g_rel <- N_g * c_g * s
T_b_rel <- N_b * c_b * s

## Species proportions
p_g <- T_g_rel / (T_g_rel + T_b_rel)
p_b <- T_b_rel / (T_g_rel + T_b_rel)

p_g
p_b

## Apply proportions to total observed tadpoles
greenfrog_tads  <- p_g * total_tads
bullfrog_tads   <- p_b * total_tads

greenfrog_tads # 3400
bullfrog_tads # 1000

## Bootstrap to get CI
B <- 5000 # sims
# bootstrap storage
p_g_boot <- numeric(B)
p_b_boot <- numeric(B)
green_tads_boot <- numeric(B)
bull_tads_boot  <- numeric(B)

for (i in 1:B) {
  
  # Bootstrap adult numbers (±10%)
  N_g_i <- rnorm(1, mean = 238, sd = 0.10 * 238)
  N_b_i <- rnorm(1, mean = 23,  sd = 0.10 * 23)
  
  # Bootstrap clutch sizes (SD = 10% of mean)
  c_g_i <- rnorm(1, mean = 5000,  sd = 0.10 * 5000)
  c_b_i <- rnorm(1, mean = 15000, sd = 0.10 * 15000)
  
  # Bootstrap survival (uniform or beta)
  s_i <- runif(1, 0.10, 0.20)   # survival between 10%–20%
  
  # Bootstrap uncertainty in total tadpole abundance
  total_tads_i <- rnorm(1, mean = 2000, sd = 0.10 * 2000)
  
  # Compute relative contributions
  T_g_rel_i <- N_g_i * c_g_i * s_i
  T_b_rel_i <- N_b_i * c_b_i * s_i
  
  # Species proportions
  p_g_boot[i] <- T_g_rel_i / (T_g_rel_i + T_b_rel_i)
  p_b_boot[i] <- 1 - p_g_boot[i]
  
  # Apply to total abundance
  green_tads_boot[i] <- p_g_boot[i] * total_tads_i
  bull_tads_boot[i]  <- p_b_boot[i] * total_tads_i
}

# Summaries
quantile(p_g_boot, c(0.025, 0.5, 0.975))
quantile(p_b_boot, c(0.025, 0.5, 0.975))

quantile(green_tads_boot, c(0.025, 0.5, 0.975))
quantile(bull_tads_boot, c(0.025, 0.5, 0.975))
bull_ci <- quantile(bull_tads_boot,  c(0.025, 0.5, 0.975))
green_ci <- quantile(green_tads_boot, c(0.025, 0.5, 0.975))

# 
# density (tadpoles per m^2)
density_est <- mean_count / pipe_area

# abundance
abundance_est <- density_est * pond_area

abundance_est

```

# Calculate true R0 for each species

## Single Species R0

### Bullfrog adults

```{r}
source("multi_sp_fxns_q1.R")
n = 200 

# set varying densities
DF_density = seq(1, 100, len = n)

movement_vals <- seq(1, 200, len = n)  # movement distances in m2

community_list <- list(
  #community1 = c("N.viridescens"),  # Newt
  #community1 = c("R.clamitans")  # Green Frog
  community1 = c("R.catesbeianus") # Bullfrog
  #community4 = c("Racl.tadpole"), # Green Frog Tadpole
 # community2 = c("Raca.tadpole")) # Bullfrog Tadpole
)

R0_res <- list()
# loop over communities
for (com in names(community_list)) {

  # pull vector of species in each community
  species_list <- community_list[[com]]
  res <- vector("list", length(DF_density) * length(movement_vals))
  # record row
  count <- 1
  # loop over zoospore densities and movement values
  for (d in DF_density) {
    for (mv in movement_vals) {
      # copy params so we don't leak changes across iterations
      ipm_d <- ipm_dists
      # set movement for all species in this community (works for 1+ species)
      for (sp in species_list) {
        ipm_d[[sp]]$movement_distance <- mv
        ipm_d[[sp]]$df_density<- d
      }
      
      # calculate R0
      R0_val <- compute_R0_stage(
          ipm_d,
          stage = "adult",
         adult_name = sp,
          tadpole_name = "Raca.tadpole", #ignores   
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150
)
      # store results as a row in a list
      res[[count]] <- data.frame(
        community = com,
        movement = mv,
        density = d,
        R0 = R0_val,
        stringsAsFactors = FALSE
      )
      count <- count + 1
    }
  }
  # bind the rows
  R0_res[[com]] <- do.call(rbind, res)
}

# bind all communities into a single df 
R0_var <- do.call(rbind, R0_res)

community_labels <- sapply(community_list, function(sp) paste(sp, collapse = " + "))

## Find Host density per liter or pond volume

per_L <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)

per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)/1000

bull_df <- R0_var %>%
  filter(is.finite(R0)) %>%
  mutate(
    community_lab = recode(as.character(community), !!!community_labels),
    host_kL = (density/per_L) * 1000,
    host_m3 = density/per_m3,
    movement_L = movement * 1000
  )

save(bull_df, file = "bull_r0_df.RData")
```

### Bull Tadpole

```{r}
source("multi_sp_fxns_q1.R")
n = 200 # change back to 50

# set varying densities
DF_density = seq(1, 800, len = n)

movement_vals <- seq(1, 20, len = n)  # movement distances in m


community_list <- list(
  #community1 = c("N.viridescens")  # Newt
  #community1 = c("R.clamitans")  # Green Frog
  #community1 = c("R.catesbeianus") # Bullfrog
  #community4 = c("Racl.tadpole"), # Green Frog Tadpole
 community1 = c("Raca.tadpole")) # Bullfrog Tadpole


R0_res <- list()
# loop over communities
for (com in names(community_list)) {

  # pull vector of species in each community
  species_list <- community_list[[com]]
  res <- vector("list", length(DF_density) * length(movement_vals))
  # record row
  count <- 1
  # loop over zoospore densities and movement values
  for (d in DF_density) {
    for (mv in movement_vals) {
      # copy params so we don't leak changes across iterations
      ipm_d <- ipm_dists
      # set movement for all species in this community 
      for (sp in species_list) {
        ipm_d[[sp]]$movement_distance <- mv
        ipm_d[[sp]]$df_density<- d
      }
      
      # calculate R0
     R0_val <- compute_R0_stage(
        ipm_d,
        stage = "tadpole",
        adult_name = "R.catesbeianus", #ignores
        tadpole_name = sp,
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150,
        include_metamorphosis_loss = TRUE
      )
      # store results as a row in a list
      res[[count]] <- data.frame(
        community = com,
        movement = mv,
        density = d,
        R0 = R0_val,
        stringsAsFactors = FALSE
      )
      count <- count + 1
    }
  }
  # bind the rows
  R0_res[[com]] <- do.call(rbind, res)
}

# bind all communities into a single df 
R0_varT <- do.call(rbind, R0_res)



community_labels <- sapply(community_list, function(sp) paste(sp, collapse = " + "))

## Find Host density per liter or pond volume

per_L <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)

per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)/1000

tad_df <- R0_varT %>%
  filter(is.finite(R0)) %>%
  mutate(
    community_lab = dplyr::recode(as.character(community), !!!community_labels),
    host_kL = (density/per_L) * 1000,
    host_m3 = density/per_m3,
    movement_L = movement * 1000
  )

save(tad_df, file = "btad_r0_df.RData")
```

### Green Frog

```{r}
source("multi_sp_fxns_q1.R")
n = 200 # change back to 50

# set varying densities
DF_density = seq(1, 800, len = n)

movement_vals <- seq(1, 200, len = n)  # movement distances in m

community_list <- list(
  #community1 = c("N.viridescens"),  # Newt
  community1 = c("R.clamitans")  # Green Frog
  #community2 = c("R.catesbeianus"), # Bullfrog
  #community4 = c("Racl.tadpole"), # Green Frog Tadpole
 # community2 = c("Raca.tadpole")) # Bullfrog Tadpole
)

R0_res <- list()
# loop over communities
for (com in names(community_list)) {

  # pull vector of species in each community
  species_list <- community_list[[com]]
  res <- vector("list", length(DF_density) * length(movement_vals))
  # record row
  count <- 1
  # loop over zoospore densities and movement values
  for (d in DF_density) {
    for (mv in movement_vals) {
      # copy params so we don't leak changes across iterations
      ipm_d <- ipm_dists
      # set movement for all species in this community
      for (sp in species_list) {
        ipm_d[[sp]]$movement_distance <- mv
        ipm_d[[sp]]$df_density<- d
      }
      adult_name <- community_list[[1]]
      # calculate R0
      R0_val <- compute_R0_stage(
          ipm_d,
          stage = "adult",
         adult_name = adult_name,
          tadpole_name = "Racl.tadpole", #ignores   
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150
)
      # store results as a row in a list
      res[[count]] <- data.frame(
        community = com,
        movement = mv,
        density = d,
        R0 = R0_val,
        stringsAsFactors = FALSE
      )
      count <- count + 1
    }
  }
  # bind the rows
  R0_res[[com]] <- do.call(rbind, res)
}


R0_varg <- do.call(rbind, res)

community_labels <- sapply(community_list, function(sp) paste(sp, collapse = " + "))

## Find Host density per liter or pond volume

per_L <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)

per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)/1000

var_df <- R0_varg %>%
  filter(is.finite(R0)) %>%
  mutate(
    community_lab = dplyr::recode(as.character(community), !!!community_labels),
    host_kL = (density/per_L) * 1000,
    host_m3 = density/per_m3,
    movement_L = movement * 1000
  )

save(var_df, file = "green_r0_df.RData")
```

### Green Frog Tadpole

```{r}
source("multi_sp_fxns_q1.R")
n = 200 # change back to 50

# set varying densities
DF_density = seq(1, 2000, len = n)

movement_vals <- seq(1, 20, len = n)  # movement distances in m


community_list <- list(
  #community1 = c("N.viridescens")  # Newt
  #community1 = c("R.clamitans")  # Green Frog
  #community1 = c("R.catesbeianus") # Bullfrog
  community4 = c("Racl.tadpole")) # Green Frog Tadpole
 #community1 = c("Raca.tadpole")) # Bullfrog Tadpole


R0_res <- list()
# loop over communities
for (com in names(community_list)) {

  # pull vector of species in each community
  species_list <- community_list[[com]]
  res <- vector("list", length(DF_density) * length(movement_vals))
  # record row
  count <- 1
  # loop over zoospore densities and movement values
  for (d in DF_density) {
    for (mv in movement_vals) {
      # copy params so we don't leak changes across iterations
      ipm_d <- ipm_dists
      # set movement for all species in this community 
      for (sp in species_list) {
        ipm_d[[sp]]$movement_distance <- mv
        ipm_d[[sp]]$df_density<- d
      }
      # calculate R0

     R0_val <- compute_R0_stage(
        ipm_d,
        stage = "tadpole",
        adult_name = "R.clamitans", #ignores
        tadpole_name = sp,
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150,
        include_metamorphosis_loss = TRUE
      )
      # store results as a row in a list
      res[[count]] <- data.frame(
        community = com,
        movement = mv,
        density = d,
        R0 = R0_val,
        stringsAsFactors = FALSE
      )
      count <- count + 1
    }
  }
  # bind the rows
  R0_res[[com]] <- do.call(rbind, res)
}

# bind all communities into a single df 
R0_var <- do.call(rbind, R0_res)

community_labels <- sapply(community_list, function(sp) paste(sp, collapse = " + "))

## Find Host density per liter or pond volume

per_L <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)

per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)/1000

gtad_df <- R0_var %>%
  filter(is.finite(R0)) %>%
  mutate(
    community_lab = recode(as.character(community), !!!community_labels),
    host_kL = (density/per_L) * 1000,
    host_m3 = density/per_m3,
    movement_L = movement * 1000
  )

save(gtad_df, file = "gtad_r0_df.RData")
```

### Newts

```{r}

source("multi_sp_fxns_q1.R")
n = 200 # change back to 50

# set varying densities
DF_density = seq(1, 1000, len = n)


movement_vals <- seq(1, 30, len = n)  # movement distances in m

community_list <- list(
  community1 = c("N.viridescens")  # Newt
  #community1 = c("R.clamitans")  # Green Frog
  #community1 = c("R.catesbeianus") # Bullfrog
  #community4 = c("Racl.tadpole"), # Green Frog Tadpole
 # community2 = c("Raca.tadpole")) # Bullfrog Tadpole
)

R0_res <- list()
# loop over communities
for (com in names(community_list)) {

  # pull vector of species in each community
  species_list <- community_list[[com]]
  res <- vector("list", length(DF_density) * length(movement_vals))
  # record row
  count <- 1
  # loop over zoospore densities and movement values
  for (d in DF_density) {
    for (mv in movement_vals) {
      # copy params so we don't leak changes across iterations
      ipm_d <- ipm_dists
      # set movement for all species in this community
      for (sp in species_list) {
        ipm_d[[sp]]$movement_distance <- mv
        ipm_d[[sp]]$df_density<- d
      }
      adult_name <- community_list[[1]]
      # calculate R0
      R0_val <- compute_R0_stage(
          ipm_d,
          stage = "adult",
         adult_name = sp,
          tadpole_name = "Racl.tadpole", #ignores   
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150
)
      # store results as a row in a list
      res[[count]] <- data.frame(
        community = com,
        movement = mv,
        density = d,
        R0 = R0_val,
        stringsAsFactors = FALSE
      )
      count <- count + 1
    }
  }
  # bind the rows
  R0_res[[com]] <- do.call(rbind, res)
}

# bind all communities into a single df 
R0_varN <- do.call(rbind, R0_res)

community_labels <- sapply(community_list, function(sp) paste(sp, collapse = " + "))

## Find Host density per liter or pond volume

per_L <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)

per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5)/1000

newt_df <- R0_varN %>%
  filter(is.finite(R0)) %>%
  mutate(
    community_lab = recode(as.character(community), !!!community_labels),
    host_kL = (density/per_L) * 1000,
    host_m3 = density/per_m3,
    movement_L = movement * 1000
  )
save(newt_df, file = "newt_r0_df.RData")
```


#Multi-stage Bullfrog R0
```{r}
source("multi_sp_fxns_q1.R")

n <- 200
Ntot_vals <- seq(1, 750 , len = n) # total bullfrog abundance in pond
meff_vals <- seq(1, 100, len = n) # effective movemetn
k_vals <- seq(0.1, 5, len = n)

sp_list <- c("R.catesbeianus", "Raca.tadpole")
# baseline densities
base_N <- sapply(sp_list, function(sp) ipm_dists[[sp]]$df_density)

# proportions
p <- base_N / sum(base_N)
p
sum(p)

# choose fixed stage ratio and movement ratio 
p_adult <- unname(p[1])
mA_obs  <- 30
mT_obs  <- 6
r_move  <- mA_obs / mT_obs # adults move 1.6 x farther than tadpoles

res <- vector("list", length(Ntot_vals) * length(meff_vals))
count <- 1

for (Ntot in Ntot_vals) {
  for (k in k_vals) {

    # split total density into stage-specific densities
    N_A <- p_adult * Ntot
    N_T <- (1 - p_adult) * Ntot

    # convert effective movement -> stage-specific movements using ratio r_move
    m_T <- k * mT_obs
    m_A <- k * mA_obs

    ipm_d <- ipm_dists
    ipm_d[["R.catesbeianus"]]$df_density <- N_A
    ipm_d[["Raca.tadpole"]]$df_density  <- N_T

    ipm_d[["R.catesbeianus"]]$movement_distance <- m_A
    ipm_d[["Raca.tadpole"]]$movement_distance  <- m_T

    R0_val <- compute_bullfrog_2stage_R0_current(
      ipm_d = ipm_d,
      adult_name = "R.catesbeianus",
      tadpole_name = "Raca.tadpole",
      pond_diameter = 22,
      lower = -15, upper = 18, bins = 150
    )

    res[[count]] <- data.frame(
      Ntot = Ntot,
      p_adult = p_adult,
      N_adult = N_A,
      N_tad   = N_T,
      k = k,
      m_adult = m_A,
      m_tad   = m_T,
      R0 = R0_val
    )
    count <- count + 1
  }
}

R0_var <- do.call(rbind, res)

# observed counts
N_A <- 23
N_T <- 450

# pond volume
per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5) / 1000

# observed total density (y-axis value)
obs_total_density_m3 <- (N_A + N_T) / per_m3
max_density_m3 <- 720/per_m3
min_density_m3 <- 310/per_m3

# observed movements (same units as x-axis)
m_A <- 30   # adult
m_T <- 6   # tadpole

# density-weighted average movement
obs_mv <- (N_A * m_A + N_T * m_T) / (N_A + N_T)

bf_df <- R0_var %>%
  mutate(host_m3_total = Ntot / per_m3) %>%
  filter(is.finite(R0))

save(bf_df, file = "multi_bf_r0_df.RData")

```



# Multi-stage green frog
```{r}

source("multi_sp_fxns_q1.R")
 
n <- 200
Ntot_vals <- seq(100, 2800, len = n)        # total bullfrog abundance in pond
#meff_vals <- seq(1, 200, len = n)  
k_vals <- seq(0.1, 5, len = n)
 
sp_list <- c("R.clamitans", "Racl.tadpole")
 # baseline densities
base_N <- sapply(sp_list, function(sp) ipm_dists[[sp]]$df_density)
 
# proportions
p <- base_N / sum(base_N)
p

sum(p)

 
# choose fixed stage ratio and movement ratio
 p_adult <- unname(p[1])
mA_obs  <- 40
mT_obs  <- 3
res <- vector("list", length(Ntot_vals) * length(meff_vals))
count <- 1

for (Ntot in Ntot_vals) {
 for (k in k_vals) {

   # split total density into stage-specific densities
   N_A <- p_adult * Ntot
   N_T <- (1 - p_adult) * Ntot

   # convert effective movement -> stage-specific movements using ratio r_move
   m_T <- mT_obs * k
   m_A <- mA_obs * k

   ipm_d <- ipm_dists
   ipm_d[["R.clamitans"]]$df_density <- N_A
   ipm_d[["Racl.tadpole"]]$df_density  <- N_T

   ipm_d[["R.clamitans"]]$movement_distance <- m_A
   ipm_d[["Racl.tadpole"]]$movement_distance  <- m_T
   
   R0_val <- compute_bullfrog_2stage_R0_current(
     ipm_d = ipm_d,
     adult_name = "R.clamitans",
     tadpole_name = "Racl.tadpole",
     pond_diameter = 22,
     lower = -15, upper = 18, bins = 150
   )

   res[[count]] <- data.frame(
     Ntot = Ntot,
     p_adult = p_adult,
     N_adult = N_A,
     N_tad   = N_T,
     k = k,
     m_adult = m_A,
     m_tad   = m_T,
     R0 = R0_val
   )
   count <- count + 1
 }
}

 R0_varg <- do.call(rbind, res)

# observed counts
N_A<- ipm_dists[["R.clamitans"]]$df_density
N_T <- ipm_dists[["Racl.tadpole"]]$df_density

# pond volume
per_m3 <- calculate_pond_volume(diameter = 22, avg_depth = 0.5) / 1000

# observed total density (y-axis value)
obs_total_density_m3 <- (N_A + N_T) / per_m3
max_density_m3 <- 2550/per_m3
min_density_m3 <- 1300/per_m3

# observed movements (same units as x-axis)
m_A <- ipm_dists[["R.clamitans"]]$movement_distance   # adult
m_T <- ipm_dists[["Racl.tadpole"]]$movement_distance   # tadpole

# density-weighted average movement
obs_mv <- (N_A * m_A + N_T * m_T) / (N_A + N_T)

gf_df <- R0_varg %>%
  mutate(host_m3_total = Ntot / per_m3) %>%
  filter(is.finite(R0))

save(gf_df, file = "multi_gf_r0_df.RData")
```
# All species

```{r}

source("multi_sp_fxns_q1.R")

n <- 200
Ntot_vals <- seq(100, 3400, len = n)  
k_vals <- seq(0.1, 5, len = n)

# calculate density ratio
sp_comm <- c("R.clamitans", "Racl.tadpole", "R.catesbeianus", "Raca.tadpole")

# baseline densities
base_N <- sapply(sp_comm, function(sp) ipm_dists[[sp]]$df_density)

# proportions
p <- base_N / sum(base_N)
p
sum(p) 

## Distance moved
mAg_obs  <- ipm_dists[["R.clamitans"]]$movement_distance
mTg_obs  <- ipm_dists[["Racl.tadpole"]]$movement_distance
mAb_obs  <- ipm_dists[["R.catesbeianus"]]$movement_distance
mTb_obs  <- ipm_dists[["Raca.tadpole"]]$movement_distance


res <- vector("list", length(Ntot_vals) * length(k_vals))
count <- 1

for (Ntot in Ntot_vals) {
  for (k in k_vals) {

    # split total density into stage-specific densities
    N_Ag <- unname(p[1]) * Ntot
    N_Tg <- unname(p[2]) * Ntot
    
    N_Ab <- unname(p[3]) * Ntot
    N_Tb <- unname(p[4]) * Ntot

    # scale df density
    ipm_d <- ipm_dists
    ipm_d[["R.clamitans"]]$df_density <- N_Ag
    ipm_d[["Racl.tadpole"]]$df_density  <- N_Tg
    ipm_d[["R.catesbeianus"]]$df_density <- N_Ab
    ipm_d[["Raca.tadpole"]]$df_density  <- N_Tb
    
    # scale movemetn
    ipm_d[["R.clamitans"]]$movement_distance <- mAg_obs * k
    ipm_d[["Racl.tadpole"]]$movement_distance  <- mTg_obs * k
    ipm_d[["R.catesbeianus"]]$movement_distance <- mAb_obs * k
    ipm_d[["Raca.tadpole"]]$movement_distance <- mTb_obs * k

    R0_val <- compute_R0_stage_community(
        species_list = sp_comm,
        ipm_dists = ipm_d,
        pond_diameter = 22,
        lower = -15, upper = 18, bins = 150,
        include_emigration = TRUE,
        include_infect_then_meta = TRUE,
        z_decay = 0.3
        )
    
    res[[count]] <- data.frame(
      Ntot = Ntot,
      N_green = N_Ag,
      N_gtad   = N_Tg,
      N_bull = N_Ab,
      N_btad   = N_Tb,
      k = k,
      m_Ag = mAg_obs * k,
      m_tg = mTg_obs * k,
      m_Ab = mAb_obs * k,
      m_Tb = mTb_obs * k,
      R0 = R0_val
    )
    count <- count + 1
  }
}

# bind all communities into a single df 
R0_var_all <- do.call(rbind, res)

save(R0_var_all, file = "R0_all_df.RData")

```


## Sensitivity Analysis


# Functions for elasticity

```{r}
setwd("C:/Users/demar/OneDrive/Documents/UTK/Research/Projects/multi_sp_competence")

source("multi_sp_fxns_q1.R")

ipm_dists <- list(R.clamitans = green_dist, R.catesbeianus = bull_dist, Raca.tadpole = btad_dist, N.viridescens = newt_dist)


species_list <- c("R.clamitans", "R.catesbeianus", "Raca.tadpole", "N.viridescens")

sens_elas <- calculate_R0_elasticity(ipm_dists, species_list, delta = 0.01)

```

### Plot the elasticities and sensitivities

```{r}
library(tidyr)
df_sens_elas <- sens_elas %>%
  pivot_longer(cols = c(sensitivity, elasticity),
               names_to = "metric",
               values_to = "value")


param_math <- c(
  growth_int   = "G[0]",
  growth_slope = "G[1]",
  lambda_int   = "lambda[0]",
  lambda_slope = "lambda[1]",
  loss_int     = "L[0]",
  loss_slope   = "L[1]",
  trans_int    = "tau[0]",
  trans_slope  = "tau[1]",
  trans_exp    = "phi",
  init_inf_int = "eta[0]",
  init_inf_sigma = "sigma[eta]"
)

## convert ready labels
sens_elas_labeled <- sens_elas %>%
  mutate(param_label_math = param_math[parameter])

# sensitivity plot

ggplot(sens_elas %>% arrange(sensitivity) +
       aes(x = reorder(parameter, sensitivity),
           y = sensitivity,
           fill = species)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(x = "Parameter",
       y = "Sensitivity") +
  theme_minimal(base_size = 14)

# Elasticity plot

elasticity_plot <- ggplot(sens_elas_labeled %>% arrange(elasticity),
       aes(x = reorder(param_label_math, elasticity),
           y = elasticity,
           fill = species)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_x_discrete(labels = function(x) parse(text = x)) +
  labs(x = "Parameter",
       y = expression("Elasticity (% change in " * R[0] * " per 1% change in parameter)")) +
  theme_minimal(base_size = 14) +
    scale_fill_manual(values = c("R.catesbeianus" = "tan4", 
                                "R.clamitans" = "forestgreen", 
                                "Raca.tadpole" = "aquamarine3",
                                "N.viridescens" = "darkorange3"))

ggsave(elasticity_plot, filename = "plots/R0_elasticity.png", width = 10, height = 6)

```


# ZOOSPORE SURVIVAL CALCULATION
```{r}

# Woodhams et al. 2008: zoospore death hazards dZ (per hour)
dZ_23  <- 1.6e-2   # 23.0 C
dZ_145 <- 3.5e-3   # 14.5 C

dt_days  <- 6
dt_hours <- dt_days * 24

#convert continuous-time hazard -> discrete-time probability over dt
hazard_to_prob <- function(h, dt_hours) 1 - exp(-h * dt_hours)

# linear interpolation at 21C
T_target <- 21.0
T1 <- 14.5
T2 <- 23.0

dZ_21 <- dZ_145 + (dZ_23 - dZ_145) *
  (T_target - T1) / (T2 - T1)

pdeath_21 <- hazard_to_prob(dZ_21, dt_hours)

out <- data.frame(
  temp_C = c(23.0, 21.0, 14.5),
  dZ_per_hour = c(dZ_23, dZ_21, dZ_145),
  p_death_6d = c(
    hazard_to_prob(dZ_23, dt_hours),
    pdeath_21,
    hazard_to_prob(dZ_145, dt_hours)
  ),
  p_survive_6d = 1 - c(
    hazard_to_prob(dZ_23, dt_hours),
    pdeath_21,
    hazard_to_prob(dZ_145, dt_hours)
  )
)

print(out)



```


